---
title: "Status of childhood immunization"
date: "September 2025"
date-format: "MMMM YYYY"
format:
   typst:
      template-partials:
         - typst-template.typ
         - typst-show.typ
      keep-typ: true
params:
   state: utah
execute:
  echo: false
  message: false
  warning: false
  fig-align: center
---

```{r}
library(tidyverse)
library(maps)
library(glue)

source("charts.R")

state <- params$state
state_title <- str_to_title(state)
hey <- "heyyy"

mock_maplot <- function() {
    world <- map_data("world")

    ggplot(world, aes(long, lat, group = group)) +
        geom_polygon(fill = "#d9d9d9", color = "#2c3e50", size = 0.3) +
        coord_fixed(1.3) +
        theme_void()
}

mock_barplot <- function() {
    df <- data.frame(
        category = c("A", "B", "C", "D"),
        value = c(4, 7, 3, 6)
    )

    ggplot(df, aes(x = value, y = reorder(category, value))) +
        geom_col(fill = "#2c3e50", width = 0.6) +
        theme_minimal(base_size = 14) +
        xlab("") +
        ylab("")
}

mock_lineplot <- function() {
    df <- data.frame(
        x = 1:10,
        y = c(3, 5, 4, 6, 8, 7, 9, 8, 10, 12)
    )

    ggplot(df, aes(x = x, y = y)) +
        geom_line(color = "#2c3e50", size = 1) +
        geom_point(color = "#2c3e50", size = 5) +
        theme_minimal(base_size = 14) +
        xlab("") +
        ylab("")
}
```

This brief illustrates the current status of childhood immunization in Maryland and is intended to inform state-level policy decisions and priorities. Takeaways include: 

- State-level immunization coverage in Maryland for key childhood vaccines, such as DTaP and MMR, is above national averages. 
- Maryland's non-medical exemption rate among kindergartners is lower than the U.S. median.
- State-level per capita public health spending is higher than the national average. 
- Maryland has had relatively low levels of measles transmission in 2025.

```{=typst} 
#blueline()
```

## Vaccination coverage

```{=typst}
#v(3pt)
```

Maintaining sufficient vaccination coverage is critical for establishing community protection. The charts below demonstrate how coverage for two critical vaccines has changed over time in Maryland and how it compares to neighboring states, national averages, and Healthy People 2030 (HP2030) targets.

<div style="background-color: #F8F8F8;">

```{=typst}
#chart-title("Proportion of 2-year-olds fully protected with DTaP vaccines")
```

::: {layout="[[1,1]]"}

```{r}
mmr_vaccination_over_time_chart_line(state_title)
```

```{r}
mock_barplot() + ggtitle("Vaccination comparison (2023)")
```

:::

In 2023, fewer two-year-olds were fully vaccinated against diphtheria, tetanus, and pertussis (received all four doses of DTaP) in Maryland compared to the previous year. Maryland has higher coverage than the US median and higher coverage than neighboring states. Coverage in Maryland is below the HP2030 target of 90%.

```{=typst}
#source("Source: ChildVaxView")
#v(7pt)
```

</div>

```{=typst} 
#pagebreak() // force pagebreak here
```

<div style="background-color: #F8F8F8;">

```{=typst}
#chart-title("Proportion of kindergartners protected with MMR vaccines")
```

::: {layout="[[1,1]]"}

```{r}
mock_lineplot() +
    ggtitle(glue("Vaccination in {str_to_title(state)} over time"))
```


```{r}
mock_barplot() + ggtitle("Vaccination comparison (2023)")
```

:::

In 2023, the same proportion of kindergarteners received at least 1 dose of measles, mumps, and rubella (MMR) vaccine in Maryland compared to the previous year. Maryland has higher coverage than the US median and higher coverage than neighboring states. Coverage in Maryland exceeds the HP2030 target of 95%.

```{=typst}
#source("Source: SchoolVaxView")
#v(7pt)
```

</div>

```{=typst} 
#blueline()
```

## Vaccination exemptions

```{=typst}
#v(3pt)
```

Many states allow children attending public school to receive vaccination exemptions for religious reasons or for personal reasons, sometimes referred to as “philosophical exemptions.” Higher rates of non-medical exemptions have been linked with increased disease transmission.

<div style="background-color: #F8F8F8;">

```{=typst}
#v(7pt)
```

```{r}
first_box <- "NO"
second_box <- "YES"
third_box <- "2%"
fourth_box <- "4%"
```

```{=typst}
#box(
  inset: 1em,
  [
    #grid(
      columns: (1fr, 1fr, 1fr, 1fr),
      gutter: 1em,
      [#status-boxes(top-text: "Allows personal exemptions?", bottom-text: "`r first_box`")],
      [#status-boxes(top-text: "Allows religious exemptions?", bottom-text: "`r second_box`")],
      [#status-boxes(top-text: "Non-medical exemptions — state", bottom-text: "`r third_box`")],
      [#status-boxes(top-text: "Non-medical exemptions — US median", bottom-text: "`r fourth_box`")]
    )
  ]
)
```

Maryland's rate of non-medical exemptions among kindergartners during the 2024-25 school year is a slight increase from the state's 2023-24 rate.

**Additional details on exemption requirements:**

A child's parent or guardian may request an exemption in writing for medical or religious  reasons. The exemption does not apply during an emergency or epidemic of disease.

```{=typst}
#source("Source: NCSL, SchoolVaxView")
#v(7pt)
```

</div>

```{=typst} 
#blueline()
```

## PUBLIC HEALTH SPENDING

<div style="background-color: #F8F8F8;">

```{=typst}
#v(7pt)
```

```{r}
left_box <- "$210"
right_box <- "$124"
```

::: {layout="[[5, 2, 1]]"}

Low levels of public health spending are thought to contribute to suboptimal immunization rates. Nationally, Maryland ranks 4th in public health spending.

```{=typst}
#text("State")
#h(0.7in)
#text("National")
#connected-boxes(text1: "`r left_box`", text2: "`r right_box`")
```

:::

```{=typst}
#source("Source: America's Health Rankings")
#v(7pt)
```

</div>

```{=typst} 
#blueline()
```

## UNIVERSAL VACCINE PURCHASING

<div style="background-color: #F8F8F8;">

```{=typst}
#v(7pt)
```

::: {layout="[[3,1]]"}

In states with Universal Purchase programs, the state government purchases all recommended vaccines for all children, regardless of insurance status. These initiatives can help to address disparities in vaccine coverage and support equitable vaccine access.

```{=typst}
#status-boxes(top-text: "Universal Purchaser?", bottom-text: "`r fourth_box`")
```

:::

```{=typst}
#source("Source: AIM")
#v(7pt)
```

</div>

```{=typst} 
#blueline()
```

## SUPPORT FOR IMMUNIZATION

Strong policy commitment to immunization is critical for effective vaccination programs. The state legislature has recently introduced several bills that would affect state-wide childhood vaccination, a selection of which are described below. The arrows below indicate whether these bills would strengthen (↑) or weaken (↓) vaccine safety nets.

<div style="background-color: #F8F8F8;">

```{=typst}
#v(7pt)
```

- HB1315 (enacted) – Broadens pharmacists' immunization roles and mandates health insurers cover all CDC-recommended childhood vaccinations.

- HB974 (enacted): Authorizes the health commissioner to establish regulations requiring health insurers to cover all vaccines recommended by the Advisory Committee on Immunization Practices (ACIP)

- SB821/HB608 (proposed) – Would extend the temporary school admission period for students lacking required immunizations from 20 to 60 days.

```{=typst}
#source("Source: AIM, LegiScan")
#v(7pt)
```

</div>

```{=typst} 
#blueline()
```

## DISEASE STATUS

Measles outbreaks can indicate insufficient vaccination coverage within a population. Disease may spread across state borders when vaccine coverage is low. The map below visualizes the number of measles cases reported in Maryland and neighboring states between January 1, 2025, and June 27, 2025. 

<div style="background-color: #F8F8F8;">

```{r}
#| layout: [[10]]
#| fig-height: 1.5
#| fig-width: 4
mock_maplot()
```

</div>

<div style="background-color: #002D72; color: white;">

```{=typst} 
#box(
  inset: 1em,
  [
    #grid(
      columns: (1fr, 10fr),
      gutter: 0.1em,
      [#image("assets/coins.svg")],
      [#text(font: "Gentona", size: 11pt)[Vaccines can help prevent expensive disease outbreaks. A 2018–19 measles outbreak in Washington was estimated to cost #text(fill: rgb("#f1c400"))[US\$47,479] per case for both direct medical and public health response expenses.]]
    )
    #source("Source: Pike, 2022")
  ]
)
```

</div>

```{=typst} 
#blueline()
```

## DATA SOURCES

Vaccination coverage:

- DTaP: CDC, ChildVaxView Interactive! https://www.cdc.gov/childvaxview/about/interactive-reports.html. DTaP, ≥ 4 Doses, States/Local Areas, Birth Years/Cohorts 2017–2021, Age 24 months. Updated Aug 2024. 
MMR: CDC, SchoolVaxView Interactive! https://www.cdc.gov/schoolvaxview/data/index.html. MMR, States, School Years 2021-22, 2022-23, 2023-24, 2024-25. Updated July 2025.

Vaccination exemptions:

- Status: NCSL, State Non-Medical Exemptions From School Immunization Requirements. https://www.ncsl.org/health/state-non-medical-exemptions-from-school-immunization-requirements. Updated July 2025.
Rates: CDC, SchoolVaxView Interactive! https://www.cdc.gov/schoolvaxview/data/index.html. Exemption – Non-Medical Exemption, States, School Years 2023-24 and 2024-25. Updated July 2025.

Public health spending:

- Public Health Funding in United States, America's Health Rankings, United Health Foundation. https://www.americashealthrankings.org/explore/measures/PH_funding. Accessed July 2025.

Universal vaccine purchasing:

- Association of Immunization Managers, Policy Maps – Universal Vaccine Purchase Program. https://www.immunizationmanagers.org/resources/aim-policy-maps/. Updated April 2025.

Support for immunization:

- Association of Immunization Managers, Legislative Round-ups. https://www.immunizationmanagers.org/resources-toolkits/immunization-program-policy-toolkit/legislative-round-ups/. 
LegiScan. https://legiscan.com/. Accessed July 2025.

Disease status:

- International Vaccine Access Center, U.S. Measles Tracker. https://publichealth.jhu.edu/ivac/resources/us-measles-tracker. Accessed July 22, 2025.

Measles outbreak cost:

- Pike J, Melnick A, Gastañaduy PA, et al. Societal Costs of a Measles Outbreak. Pediatrics. 2021;147(4):e2020027037. doi:10.1542/peds.2020-027037

*Note: The high-level data included in this report do not reflect statewide variation in vaccination coverage or disease status. Further, state reporting policies may limit data completeness. For any data-related questions, please contact ivac@jh.edu.*
